---
title: Ana-Maria 1.4 
output: html_notebook
---
[BIEN](http://biendata.org)
```{r}
Zanthoxylum_americanum <- BIEN_occurrence_species(species = "Zanthoxylum americanum")
str(Zanthoxylum_americanum)
head(Zanthoxylum_americanum)
tail(Zanthoxylum_americanum)
```
Information for my species has been downloaded so now we will use this to make a rangemap
```{r}
zanthoxylum_americanum_rangemap <- BIEN_ranges_load_species("Zanthoxylum_americanum")

# Mapping the range of zanthoxylum americanum 
map('world', fill=T, col="grey", bg = "light blue", xlim=c(-180,20),ylim=c(-60,80))
plot(zanthoxylum_americanum_rangemap, col="green", add=T)
```
```{r}
#Now we will map the range and occurence of pricklyash
map('world', fill=T, col="grey", bg="light blue", xlim=c(-180,-20), ylim=c(-60, 80))
plot(zanthoxylum_americanum_rangemap, col="green", add=T)
points(cbind(Zanthoxylum_americanum$longitude,Zanthoxylum_americanum$latitude), col="blue", pch=20, cex=1)
```
```{r}
#Now that we have the range and occurence for pricklyash we will expand this data set for the entire world 
Zanthoxylum_americanum_full <- BIEN_occurrence_species(species = "Zanthoxylum americanum",cultivated = T,only.new.world = F,all.taxonomy = T,native.status = T,observation.type = T,political.boundaries = T)

str(Zanthoxylum_americanum_full)
```
```{r} 
#now we will map this 
map('world', fill=T , col="grey", bg="light blue")

#here we are plotting ALL points in red
points(cbind(Zanthoxylum_americanum_full$longitude, Zanthoxylum_americanum_full$latitude), col="red", pch=20, cex=1)

#here we are plotting points from oroginal data set in blue
points(cbind(Zanthoxylum_americanum$longitude, Zanthoxylum_americanum$latitude), col="blue", pch=20, cex=1) 
```
```{r}
#now we will map another a cousin of pricklyash, known as wid lime
Zanthoxylum_fagara <- BIEN_occurrence_species(species = "Zanthoxylum fagara")
str(Zanthoxylum_fagara)
head(Zanthoxylum_fagara)
tail(Zanthoxylum_fagara)
```

```{r}
#now we will make a range map for wild lime
zanthoxylum_fagara_rangemap <- BIEN_ranges_load_species("Zanthoxylum_fagara")


map('world', fill=T, col="grey", bg = "light blue", xlim=c(-180,20),ylim=c(-60,80))
plot(zanthoxylum_fagara_rangemap, col="green", add=T)
```
```{r}
#now we will map range and occurence for wild lime 
map('world', fill=T, col="grey", bg="light blue", xlim=c(-180,-20), ylim=c(-60, 80))
plot(zanthoxylum_fagara_rangemap, col="green", add=T)
points(cbind(Zanthoxylum_fagara$longitude,Zanthoxylum_fagara$latitude), col="blue", pch=20, cex=1)
```
```{r}
#now we will expand this same data set for the world
Zanthoxylum_fagara_full <- BIEN_occurrence_species(species = "Zanthoxylum fagara",cultivated = T,only.new.world = F,all.taxonomy = T,native.status = T,observation.type = T,political.boundaries = T)

str(Zanthoxylum_fagara_full)
```
```{r}
#now we will map the added data with out original data set of wild lime 
map('world', fill=T , col="grey", bg="light blue")

#here we are plotting ALL points in red
points(cbind(Zanthoxylum_fagara_full$longitude, Zanthoxylum_fagara_full$latitude), col="red", pch=20, cex=1)

#here we are plotting points from oroginal data set in blue
points(cbind(Zanthoxylum_fagara$longitude, Zanthoxylum_fagara$latitude), col="blue", pch=20, cex=1) 
```
```{r}
#Now we will begin to do a statistical analysis of one trait between these two species
BIEN_trait_list()
```
```{r}
#now we will create a frame for Zanthoxylum traits 
Zanthoxylum_traits <- BIEN_trait_genus(genus = "Zanthoxylum")
str(Zanthoxylum_traits)
Zanthoxylum_traits
```
```{r}
#Now we will choose a trait, I chose seed mass 
zanthoxylum_sm <- BIEN_trait_traitbygenus(trait = "seed mass", genus = "Zanthoxylum")
head(zanthoxylum_sm)
tail(zanthoxylum_sm)
str(zanthoxylum_sm)
```
```{r}
#Now we will list the variables from the data frame
names(zanthoxylum_sm)
```
```{r}
zanthoxylum_sm_small = zanthoxylum_sm[c(1:5)]
zanthoxylum_sm_small
zanthoxylum_sm_small = as.data.frame(zanthoxylum_sm_small)
class(zanthoxylum_sm_small)
head(zanthoxylum_sm_small)
class(zanthoxylum_sm_small$trait_value)
sm = as.numeric(zanthoxylum_sm_small$trait_value)
class(sm)
```
```{r}
#now we will use this data to make a box plot showing the data we gathered about seed mass
library(lattice)
head(zanthoxylum_sm_small)
boxplot(sm~scrubbed_species_binomial, data=zanthoxylum_sm_small)
```
```{r}
#now I will pick another trait for the same genus, in this case I picked whole plant height
zanthoxylum_ph <- BIEN_trait_traitbygenus(trait = "whole plant height", genus = "Zanthoxylum")
head(zanthoxylum_ph)
tail(zanthoxylum_ph)
str(zanthoxylum_ph)
```
```{r}
#listing the variables from the data frame 
names(zanthoxylum_ph)

```
```{r}
#now we wil continue to define the variable 
zanthoxylum_ph_small =zanthoxylum_ph[c(1:5)]
zanthoxylum_ph_small
zanthoxylum_ph_small = as.data.frame(zanthoxylum_ph_small)
class(zanthoxylum_ph_small)
head(zanthoxylum_ph_small)
class(zanthoxylum_ph_small$trait_value)
ph = as.numeric(zanthoxylum_ph_small$trait_value)
class(ph)
```
```{r}
#now we will make a box plot to analyze the data on whole plant height 
library(lattice)
head(zanthoxylum_ph_small)
boxplot(ph~scrubbed_species_binomial, data=zanthoxylum_ph_small)
```
```{r}
#now we will summarize the data, first for plant height and then for seed mass 
mean(ph, data=zanthoxylum_ph_small)
mean(sm, data=zanthoxylum_sm_small)
```

In this assignemnt I learned how to map both range and occurence for plant species and this was shown through the two maps that I did mapping prickly ash and wild lime, both "cousins" within the Zanthoxylum family. Afterwards I decided to go within the general Genus of of the Zanthoxylum family to compare traits. From this genus I chose to compare seed mass and whole plant height. First I created a box plot for both in order to gange a general summary of the distribution for seed mass across two species of zanthoxylum, and then I did the same, except for plant heght. Afterwards I conducted a summary statistics and gathered that the mean seed mass for zanthoxylum in my designated data range was 19.29608mg and the mean whole plant height within the data range I used was 10.467meters. To complete this assigment I will do a histgram showng the distribution of plant height. 
```{r}
#box plot 
boxplot(ph)
histogram(ph)

```
```{r}
# attempting a test 
ttest_result1 = t.test(x=zanthoxylum_ph_small$trait_value, y=zanthoxylum_sm_small$trait_value, paired=FALSE, alternative="two.sided")
ttest_result1
```
#Here I tried to finish the portfolio by doing an unpaired t test between two characteristics, however I wasn't able to get the code right. I need to continue working on this 

